---
title: "Motor-unit decomposition of HD-sEMG"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Motor-unit decomposition of HD-sEMG}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

`PhysioHDEMG` decomposes high-density surface EMG (HD-sEMG) electrode-grid
recordings into the discharge patterns of the individual motor units that
produced them. This vignette runs the whole pipeline offline on a synthetic
recording with **known** motor units, so every number can be checked against
ground truth — no device or data file is needed.

```{r}
library(PhysioHDEMG)
```

## 1. Simulate a recording with known motor units

`make_hdemg_sim()` superimposes a set of motor units — each with its own
discharge train and its own spatial action-potential map over the electrode
grid — and returns the ground-truth discharge times alongside the signal.

```{r}
sim <- make_hdemg_sim(
  n_units      = 3,
  duration_sec = 4,
  grid         = c(4, 4),   # 16-channel electrode grid
  noise_sd     = 0.005,     # smaller = higher SNR
  seed         = 1
)
sim
dim(sim$signal)             # n_time x n_channels
lengths(sim$spikes)         # ground-truth discharges per unit
```

## 2. Decompose

`hdEMGDecompose()` extends and whitens the channels, seeds each source by a
fixed-point contrast maximisation, refines it by convolution-kernel
compensation, and accepts units that clear a pulse-to-noise-ratio and
silhouette quality gate.

```{r}
dec <- hdEMGDecompose(sim, n_units = 6, extension_factor = 10, seed = 1)
dec
dec$quality
```

## 3. Inspect the recovered units

The accessors turn a decomposition into tidy tables. `muPulseTrains()` lists
every discharge, `muFiringStats()` summarises discharge rate, recruitment, and
quality per unit, and `muActionPotentialMaps()` recovers each unit's spatial
footprint on the grid.

```{r}
head(muPulseTrains(dec))
muFiringStats(dec)

maps <- muActionPotentialMaps(dec)
maps[[1]]$peak_channel     # strongest electrode for the first unit
```

## 4. Validate against the ground truth

Because the simulator recorded the true discharge times, we can score the
decomposition with a lag-aware rate of agreement. `roa_lag_aware` near 1 means
a reference unit was recovered; `lag` is the constant latency between the two.

```{r}
matchMotorUnits(dec, sim$spikes, tol = 2, max_lag = 5)
```

## Working with your own recordings

On real data, read a grid from an HDF5 file with `readHDEMG()` (which can also
carry a bundled reference decomposition for `matchMotorUnits()`), or wrap an
in-memory matrix with `hdemgToPhysioExperiment()` to join the shared
`PhysioExperiment` data model. Real motor units have lower quality than this
clean simulation, so band-pass the signal (`bandpass = c(20, 500)`) and relax
the gates (`pnr_threshold`, `sil_threshold`) — see `?hdEMGDecompose`.

Conventional (non-decomposed) EMG features live in the sibling `PhysioEMG`
package.
